---
title: "Automatic Regression Modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Automatic Regression Modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA
)
```

## Installation

You can install autoReg package on github. 

```{r, eval=FALSE}
#install.packages("devtools")
devtools::install_github("cardiomoon/autoReg")
```

## Load package

To load the package, use library() function.
```{r}
library(autoReg)
library(dplyr)
```

## Linear model with multiple variables

The package `autoReg` aims automatic selection of explanatory variables of regression models. Let's begin with famous mtcars data. We select mpg(miles per gallon) as a dependent variable and select wt(weight), hp(horse power) and am(transmission, 0=automatic, 1=manual) as explanatory variables and included all possible interaction. The autoReg() function make a table summarizing the result of analysis.


```{r}
fit=lm(mpg~wt*hp*am,data=mtcars)
autoReg(fit) 
```

You can make a publication-ready table easily using myft(). It makes a flextable object which can use in HTML, PDF, microsoft word or powerpoint file.

```{r}
autoReg(fit) %>% myft()
```

From the result of multivariable analysis, we found no explanatory variable is significant.

### Selection of explanatory variable from univariable model

You can staret with univariable model. With a list of univariable model, you can select potentially significant explanatory variable(p value below 0.2 for example). The autoReg() function automatically select from univariable model with a given p value threshold(default value is 0.2).

```{r}
autoReg(fit,uni=TRUE, threshold=0.2) %>% myft()
```
As you can see in the above table, the coefficents of hp:am(the interaction of hp and am) and wt:hp:am (interaction of wt and hp and am) have p-values above 0.2. So these variables are excluded and the remaining variables are used in multivariable model.

### Stepwise backward elimination  

From the multivariable model, you can perform stepwise backward elimination with step() function.

```{r}
fit=lm(mpg~wt+hp+am+wt:hp+wt:am,data=mtcars)
final=step(fit,trace=0)
summary(final)
```
You can perform univarible anlaysis for variable selection, mutivariable analysis and stepwise backward elimination in one step.

```{r}
fit=lm(mpg~hp*wt*am,data=mtcars)
autoReg(fit,uni=TRUE,final=TRUE) %>% myft()
```

## Linear model with interaction between categorical variable

You can use autoReg() function for models with interaction with categorical variable(s).

```{r}
fit=lm(Sepal.Length~Sepal.Width*Species,data=iris)
autoReg(fit,uni=TRUE, final=TRUE) %>% myft()
```

## Missing data - automatic multiple imputation


```{r}
library(survival)
data(cancer)
gaze(colon,show.n=TRUE) %>% myft()
colon$diff=factor(colon$diff,labels=c("well","moderate","poor"))
fit=glm(status~rx+age+nodes+diff,data=colon)
autoReg(fit,imputed=TRUE) %>% myft()
```
